\chapter{Change Log}

\section{Changes Since Proposed Final Draft}

\begin{itemize}
\item Sections \ref{request_matching} and \ref{determine_response_type}: Q-values not used in \Consumes\ or \Produces.
\item Chapter \ref{runtimedelegate}: Clarify that an implementation can supply an alternate \rd\ API class.
\end{itemize}

\section{Changes Since Public Review Draft}

\begin{itemize}
\item Chapter \ref{applications}: Renamed ApplicationConfig class to Application.
\item Chapter \ref{resources}: UriBuilder reworked to always encode components.
\item Sections \ref{resource_class_constructor} and \ref{provider_class_constructor}: Added requirement to warn when choice of constructor is ambiguous.
\item Section \ref{resource_field}: \code{FormParam} no longer required to be supported on fields or properties.
\item Section \ref{resource_method_return}: Added text describing how to determine raw and generic types from method return type and returned instance.
\item Section \ref{uritemplates}: Template parameters can specify the regular expression that forms their capturing group.
\item Section \ref{reqpreproc}: Make pre-processed URIs available rather than original request URI. Added URI normalization.
\item Section \ref{reqpreproc}: Removed URI-based content negotiation.
\item Section \ref{request_matching}: Reorganized the request matching algorithm to remove redundancy and improve readability, no functional change.
\item Section \ref{template_to_regex}: Changes to regular expressions to eliminate edge cases.
\item Section \ref{entity_providers}: Added requirement to use JavaBean Activation Framework when no entity provider can be found.
\item Section \ref{standard_entity_providers}: Require standard JAXB entity providers to use application-supplied JAXB contexts in preference to their own.
\item Section \ref{contextprovider}: Added support for specifying media type capabilities of context providers.
\item Section \ref{contexttypes}: Removed \code{ContextResolver} from list of injectable resources.
\item Section \ref{providercontext}: Changed name to Providers, removed entity provider-specific text to reflect more generic capabilities.
\item Chapter \ref{headersupport}: New appendix describing where particular HTTP headers are supported.
\end{itemize}